```{r setup, include=FALSE}
# cluster of libraries for easier R syntax
library(tidyverse)
```

# Data Visualization: Annotations

This notebook visualizes the annotation data generated in scope of this project.
The unit of analysis is per use case (not aggregated per requirement).

```{r data-loading}
d <- read.csv('data/output/uca-annotations.csv')
figdir <- 'figures/descriptive/uc/'
```

## Base Attributes

First, we visualize the different forms that are used to capture use cases.

```{r uc-form}
d %>% ggplot(aes(x=Form)) +
    geom_histogram(stat="count") + 
    geom_text(stat='count', aes(label=..count..), nudge_y=10, size=4) +
        theme(axis.title.x = element_blank(), axis.title.y = element_blank())

#ggsave(paste0(figdir, 'forms.pdf'), width=5, height=6)
```

Next, we visualize the frequency with which the different fields appear in the use cases.

```{r uc-fields}
field.order = c("title", "actor", "level", "preconditions", "postconditions", "main scenario", "extensions",
    "assumptions", "additional information", "solution hints", "dependencies", "limitations", "revision history")

d %>% 
    filter(Form == "description") %>% # focus only on use case descriptions
    select(req, ucid, Fields) %>% 
    mutate(field = strsplit(Fields, "; ")) %>% # split the Fields column into a list
    unnest(field)  %>% # "unnest" the column, which creates one entry per UC per field
    mutate(field = factor(field, levels=field.order)) %>% 
    ggplot(aes(y=field)) + # visualize the occurrence of fields in a histogram
        geom_histogram(stat="count") + 
        geom_text(stat='count', aes(label=..count..), nudge_x=10, size=4) +
        scale_y_discrete(limits=rev) +
        theme(axis.title.x = element_blank(), axis.title.y = element_blank())

#ggsave(paste0(figdir, 'fields.pdf'), width=5, height=6)
```

Addditionally, we visualize where the use cases are located.

```{r uc-location}
d %>% 
    filter(Location != "NA") %>% 
    ggplot(aes(x=Location)) +
        geom_histogram(stat="count") + 
        geom_text(stat='count', aes(label=..count..), nudge_y=10, size=4) +
        theme(axis.title.x = element_blank(), axis.title.y = element_blank())

#ggsave(paste0(figdir, 'location.pdf'), width=5, height=6)
```

The level of a use case is only rarely given.

```{r uc-level}
d %>% 
    filter(Level != "NA") %>% 
    ggplot(aes(x=Level)) +
        geom_histogram(stat="count") + 
        geom_text(stat='count', aes(label=..count..), nudge_y=10, size=4) +
        theme(axis.title.x = element_blank(), axis.title.y = element_blank())
        
#ggsave(paste0(figdir, 'level.pdf'), width=5, height=6)
```

## Scenario Attributes

Next are the attributes of the scenarios described in the use case.
First are the number and types of steps.

```{r steps}
step.order <- c("Steps", "Functional.Steps", "Whitebox.Steps", "Consistent.Grammar", "Misplaced.Variations", "Misplaced.Main")

d %>% 
    filter(Form=="description" & Procedural) %>% 
    pivot_longer(
        cols = c(Steps, Functional.Steps, Whitebox.Steps, Misplaced.Variations, Misplaced.Main, Consistent.Grammar), 
        names_to = "metric") %>% 
    mutate(metric = factor(metric, levels=step.order)) %>% 
    ggplot(aes(x=value, y=metric)) +
        geom_boxplot() +
        scale_y_discrete(labels = c("Misplaced main steps", "Misplaced variations", "Consistent grammar", "White-box steps", "Functional steps", "Steps"), limits=rev) +
        theme(axis.title.x = element_blank(), axis.title.y = element_blank())

#ggsave(paste0(figdir, 'steps.pdf'), width=4, height=3)
```

Next are the number and types of entities.

```{r entities}
entity.order <- c("entities", "actors", "actors.explicit")

d %>% 
    filter(Form=="description" & Procedural) %>% 
    pivot_longer(
        cols = c(entities, actors, actors.explicit), 
        names_to = "metric") %>% 
    mutate(metric = factor(metric, levels=entity.order)) %>% 
    ggplot(aes(x=value, y=metric)) +
        geom_boxplot() +
        scale_y_discrete(labels = c("Explicit actors", "Actors", "Entities"), limits=rev) +
        theme(axis.title.x = element_blank(), axis.title.y = element_blank())

#ggsave(paste0(figdir, 'entities.pdf'), width=4, height=1.5)
```

Finally, we visualize the distribution of interactions.
One step in a use case (i.e., one item in a scenario) can contain multiple interactions.

```{r interactions}
interaction.order <- c("interactions", "pureinteractions", "consecutive")

d  %>% 
    filter(Form=="description" & Procedural) %>% 
    pivot_longer(
        cols = c(interactions, pureinteractions, consecutive), 
        names_to = "metric") %>% 
    mutate(metric = factor(metric, levels=interaction.order)) %>% 
    ggplot(aes(x=value, y=metric)) +
        geom_boxplot() +
        scale_y_discrete(labels = c("Consecutive interactions", "User-System interactions", "Interactions"), limits=rev) +
        theme(axis.title.x = element_blank(), axis.title.y = element_blank())

#ggsave(paste0(figdir, 'interactions.pdf'), width=4, height=1.5)
```
