```{r setup, include=FALSE}
# cluster of libraries for easier R syntax
library(tidyverse)
```

# Data Visualization: Full

This notebook visualizes the annotation data generated in scope of this project.
The unit of analysis is per requirement (i.e., aggregated use cases).

```{r directory}
figdir <- 'figures/descriptive/req/'
```

## Load and Prepare Data

To begin, load the aggregated data from the `uca-aggregated.csv` file.

```{r data-loading}
d <- read.csv('data/output/uca-aggregated.csv')
d$month <- ymd(d$month)
```

Create one helper variable: `uca` (i.e., use case approach) whether a requirement contains any use case at all.

```{r detect-uca}
d <- d %>% 
    mutate(uca = ifelse(!is.na(nuc), nuc>0, 0))
```

## Visualize Data

Next, we generate descriptive statistics from the assembled data.
The statistics are categorized by certain levels that apply filters to the dataset:

1. [All requirements](#all-requirements): all requirements in the database
2. [Use Case approach](#use-case-approach): requirements containing at least one use case
3. [Use Case descriptions](#use-case-descriptions): requirements containing at least one use case description
4. [Procedural Use Case Descriptions](#procedural-uce-case-descriptions): requirements containing at least one procedural use case description

### All Requirements

We are interested in the relative number of requirements per month that contained any use case at all.

```{r uca-relative}
d %>% 
    select(month, uca) %>% 
    group_by(month) %>% 
    summarize(uc = mean(uca)) %>%  # percentage of requirements with use cases
    ggplot(aes(x=month, y=uc)) + 
        geom_line() +
        geom_smooth(method = "lm") +
        scale_y_continuous(labels = scales::percent) +
        labs(y="Fraction of requirements that contain at least one use case", x="Months")

#ggsave(paste0(figdir, 'uca-relative.pdf'), width=10, height=4)
```

Additionally, we can compare the number of requirements containing at least one use case with the number of requirements per month.

```{r uca-absolute}
d %>% 
    select(month, uca) %>% 
    group_by(month) %>% 
    summarize(
        nreq = n(), # number of requirements
        uc = sum(as.numeric(uca)) # percentage of requirements with use cases
    ) %>% 
    pivot_longer(cols=c(nreq, uc), names_to="type") %>% 
    ggplot(aes(x=month, y=value, color=type)) + 
        geom_line() + 
        geom_smooth(method = "lm") +
        scale_color_discrete(labels = c("All Requirements", "Requirements with at least one use case")) +
            theme(legend.position="bottom") +
        labs(y="Absolute number of requirements", x="Months", color="Requirement type")

#ggsave(paste0(figdir, 'uca-absolute.pdf'), width=10, height=4)
```

### Use Case approach

The following set of figures only applies to requirements that contain at least one use case (i.e., `uca==TRUE`).

```{r data-d3}
d3 <- d %>% 
    filter(uca == TRUE)
```

First, we visualize how the average number of use cases per requirement developed over time.

```{r nouc}
d3 %>%
    select(month, nuc) %>% 
    group_by(month) %>%
    summarize(n = mean(nuc)) %>% 
    ggplot(aes(x = month, y = n)) +
        geom_line() + 
        geom_smooth(method = "lm") +
        labs(y = "Average number of use cases per Requirement", y="Month")
        
#ggsave(paste0(figdir, 'nouc.pdf'), width=10, height=4)
```

Next, we want to see how the distribution of forms of use cases changed over time.

```{r forms-relative}
d3 %>%
    select(month, form) %>% 
    group_by(month) %>% 
    summarize(
        n = n(), # number of requirements
        n_desc = sum(form == "description"),
        n_title = sum(form == "title"),
        n_inf = sum(form=="informal"),
        n_us = sum(form=="user story")
    ) %>% 
    mutate(
        frac_desc = n_desc/n,
        frac_title = n_title/n,
        frac_inf = n_inf/n,
        frac_us = n_us/n
    ) %>% 
    pivot_longer(cols=c(frac_desc, frac_title, frac_inf, frac_us), names_to = 'form', values_to = "value")  %>% 
    ggplot(aes(x = month, y = value, fill = form)) +
        geom_bar(position="stack", stat="identity") +
            scale_y_continuous(labels = scales::percent) +
            scale_fill_discrete(labels = c("description", "informal", "title", "user story")) +
            theme(legend.position="bottom") +
        labs(y = "Fraction of requirements with use cases in a specific form", y="Month")
        
#ggsave(paste0(figdir, 'forms-relative.pdf'), width=10, height=4)
```

### Use Case Description

The following set of figures only applies to requirements that contain at least one use case (i.e., `uca==TRUE`) that has the form `description`.

```{r data-d4}
d4 <- d3 %>% 
    filter(form == "description")
```

Visualize the level on which use cases are specified.

```{r level}
d4 %>% 
    select(month, level) %>% 
    group_by(month) %>% 
    summarize(
        n = n(), # number of requirements
        n_sys = sum(level == "System"),
        n_bus = sum(level == "Business"),
        n_user = sum(level == "User"),
        n_no = sum(level == "(no level)")
    ) %>% 
    mutate(
        frac_sys = n_sys/n,
        frac_bus = n_bus/n,
        frac_user = n_user/n,
        frac_no = n_no/n
    ) %>% 
    pivot_longer(cols=c(frac_sys, frac_bus, frac_user, frac_no), names_to = 'level', values_to = "value")  %>% 
    ggplot(aes(x = month, y = value, fill = level)) +
        geom_bar(position="stack", stat="identity") +
            scale_y_continuous(labels = scales::percent) +
            scale_fill_discrete(labels = c("business", "(no level)", "system", "user")) +
            theme(legend.position="bottom") +
        labs(y = "Fraction of requirements with use case descriptions on a specific level", y="Month")
            
#ggsave(paste0(figdir, 'level.pdf'), width=10, height=4)
```

Visualize the fraction of requirements per month that contain use case descriptions that are procedural.

```{r procedural}
d4 %>% 
    select(month, procedural) %>% 
    group_by(month) %>% 
    summarize(proc = mean(procedural)) %>% 
    ggplot(aes(x=month, y=proc)) +
        geom_line() +
        geom_smooth(method="lm") +
            scale_y_continuous(labels = scales::percent) +
        labs(y = "Fraction of requirements with procedural use case descriptions", y="Month")

#ggsave(paste0(figdir, 'procedural.pdf'), width=10, height=4)
```

### Procedural Use Case Descriptions

The following set of figures only applies to requirements that contain at least one use case (i.e., `uca==TRUE`) that has the form `description` and is `procedural`.

```{r data-d4}
d5 <- d4 %>% 
    filter(procedural > 0)
```

Visualize the proportion of requirements containing procedural use case descriptions with a proper, cogent text order.

```{r textorder}
d5 %>% 
    select(month, textorder) %>% 
    group_by(month) %>% 
    summarize(order = mean(textorder)) %>% 
    ggplot(aes(x=month, y=order)) +
        geom_line() +
        geom_smooth(method="lm") +
            scale_y_continuous(labels = scales::percent) +
        labs(y = "Fraction of requirements with a procedural use case description of proper text order", y="Month")

#ggsave(paste0(figdir, 'textorder.pdf'), width=10, height=4)
```

Visualize the proportion of requirements containing procedural use case descriptions with end-to-end dependencies.
Such requirements start with a user-to-system interaction and end with a system-to-user interaction.

```{r dependencies}
d5 %>% 
    select(month, dependencies) %>% 
    group_by(month) %>% 
    summarize(dep = mean(dependencies)) %>% 
    ggplot(aes(x=month, y=dep)) +
        geom_line() +
        geom_smooth(method="lm") +
            scale_y_continuous(labels = scales::percent) +
        labs(y = "Fraction of requirements with a procedural use case description and end-to-end dependencies", y="Month")
        
#ggsave(paste0(figdir, 'dependencies.pdf'), width=10, height=4)
```

Visualize the evolution of the average coherence.

```{r coherence}
d5 %>% 
    select(month, coherence) %>% 
    group_by(month) %>% 
    summarize(coh = mean(coherence)) %>% 
    ggplot(aes(x=month, y=coh)) +
        geom_line() +
        geom_smooth(method="lm") +
            scale_y_continuous(labels = scales::percent) +
        labs(y = "Average level of coherence", y="Month")
        
#ggsave(paste0(figdir, 'coherence.pdf'), width=10, height=4)
```

Visualize the evolution of the adherence to the correct ordering of sequences (i.e., the main success scenario and the alternative scenarios).

```{r sequence}
d5 %>% 
    select(month, sequence.main, sequence.alt, numbering) %>% 
    group_by(month) %>% 
    summarize(
        main = mean(sequence.main),
        alt = mean(sequence.alt),
        numb = mean(numbering)
    ) %>% 
    pivot_longer(cols=c(main, alt, numb), names_to = "part")  %>% 
    ggplot(aes(x=month, y=value, color=part)) +
        geom_line() +
            scale_y_continuous(labels = scales::percent) +
            scale_fill_discrete(labels = c("separate alternative scenario", "numbered main scenario", "numbered alternative scenario")) +
            theme(legend.position="bottom") +
        labs(y = "Fraction of properly numbered scenarios", y="Month")
        
#ggsave(paste0(figdir, 'sequence.pdf'), width=10, height=4)
```

#### Steps

Next, visualize the steps recorded per use case.

```{r steps}
d5 %>% 
    select(month, steps) %>% 
    group_by(month) %>% 
    summarize(
        n = n(),
        avg = mean(steps),
        var = var(steps),
    ) %>% 
    ggplot(aes(x=month, y=avg)) +
        geom_line() + 
        geom_ribbon(aes(ymin=avg-var, ymax=avg+var), alpha=0.3) +
            labs(y="Average number of steps", x="Month")
        
#ggsave(paste0(figdir, 'steps.pdf'), width=10, height=4)
```

```{r steps-all}
d5 %>% 
    select(month, steps, steps.functional, steps.mis.main, steps.mis.variations, steps.consistent) %>% 
    group_by(month) %>% 
    summarize(
        n = n(),
        step = mean(steps),
        step.fun = mean(steps.functional),
        step.mis.main = mean(steps.mis.main),
        step.mis.var = mean(steps.mis.variations),
        step.con = mean(steps.consistent)
    ) %>% 
    pivot_longer(
        cols = c(step, step.fun, step.mis.main, step.mis.var, step.con),
        names_to = 'steptype') %>% 
    ggplot(aes(x=month, y=value, color = steptype)) +
        geom_line() + 
            labs(y="Average number of steps", x="Month", color="Type of Step") +
            scale_color_discrete(labels = c("steps", "conssistent grammar", "functional steps", "misplaced main step", "misplaced variation")) +
            theme(legend.position="bottom")

#ggsave(paste0(figdir, 'steps-all.pdf'), width=10, height=4)
```

#### Interactions

On a more detailed level: visualize the number of interactions per use case.

```{r interactions}
d5 %>% 
    select(month, interactions) %>% 
    group_by(month) %>% 
    summarize(
        n = n(),
        avg = mean(interactions),
        var = var(interactions),
    ) %>% 
    ggplot(aes(x=month, y=avg)) +
        geom_line() + 
        geom_ribbon(aes(ymin=avg-var, ymax=avg+var), alpha=0.3) +
            labs(y="Average number of Interactions", x="Month")
        
#ggsave(paste0(figdir, 'interactions.pdf'), width=10, height=4)
```

```{r interactions-all}
d5 %>% 
    select(month, interactions, steps.beyond, pureinteractions, consecutive) %>% 
    group_by(month) %>% 
    summarize(
        n = n(),
        int = mean(interactions),
        int.beyond = mean(steps.beyond),
        int.pure = mean(pureinteractions),
        int.consecutive = mean(consecutive)
    ) %>% 
    pivot_longer(
        cols = c(int, int.beyond, int.pure, int.consecutive),
        names_to = 'inttype') %>% 
    ggplot(aes(x=month, y=value, color = inttype)) +
        geom_line() + 
            labs(y="Average number of interactions", x="Month", color="Type of interaction") +
            scale_color_discrete(
                labels = c(
                    "interactions",
                    "system-system interactions",
                    "consecutive interactions",
                    "system-user interactions")
            ) +
            theme(legend.position="bottom")

        
#ggsave(paste0(figdir, 'interactions-all.pdf'), width=10, height=4)
```

#### Actors

Finally, visualize the evolution of entities in the use cases.

```{r entities-all}
d5 %>% 
    select(month, entities, actors, actors_explicit) %>% 
    group_by(month) %>% 
    summarize(
        ent = mean(entities),
        act = mean(actors),
        expl = mean(actors_explicit)
    ) %>% 
    pivot_longer(
        cols = c(ent, act, expl),
        names_to = 'enttype') %>% 
    ggplot(aes(x=month, y=value, color = enttype)) +
        geom_line() + 
            labs(y="Average number of entities", x="Month", color="Type of entity") +
            scale_color_discrete(
                labels = c(
                    "actors",
                    "entities",
                    "explicit actors")
            ) +
            theme(legend.position="bottom")
        
#ggsave(paste0(figdir, 'entities.pdf'), width=10, height=4)
```
