```{r setup, include=FALSE}
# cluster of libraries for easier R syntax
library(tidyverse)
```

# Data Mocking: Annotations

Because we cannot share the real, sensitive data from this case study, we provide a script that simulates data.
This simulated data can then be used in the [visualization-annotations.Rmd](./../visualization/visualization-annotations.Rmd) script.

```{r create_requirements}
n <- 1000

reqs <- data.frame(
    req = paste0("req", 1:n), # create n requirements
    nuc = rbinom(n, 5, 0.3), # associate each requirement with up to 5 use cases
    complexity = sample(1:3, n, replace=TRUE)
)

reqs.export <- reqs %>% 
    mutate(
        cplx = ifelse(complexity == 1, "Low Complexity", 
        ifelse(complexity == 2, "Medium Complexity", "High Complexity"))
    ) %>% 
    select(req, cplx, complexity)
write.csv(reqs.export, 'data/output/reqs-mocked.csv')
```

### Use Cases

For every use case associated with a requirement, create a separate row.

```{r expand-ucs}
ucs <- reqs %>% 
    filter(nuc > 0) %>%  # focus on the requirements with at least one use case
    group_by(req) %>% 
    rowwise() %>% 
    mutate(ucid = list(paste0("uc", 1:nuc))) %>% 
    unnest(cols = c(ucid)) %>% 
    select(req, ucid, complexity)
```

Assign a `Form` to each use case with the following method

```{r select-form}
get_form <- function() {
    random <- runif(1)
    if (random < 0.5) {
        return ("description")
    } else if (random >= 0.5 && random < 0.8) {
        return ("title")
    } else if (random >=0.8 && random < 0.9) {
        return ("user story")
    } else {
        return ("informal")
    }
}

ucs <- ucs %>% 
    mutate(Form = get_form())
```

### Use Case Descriptions

For all use cases with the form `description`, determine if they are `Procedural` or not.

```{r procedural}
fields <- c("actor", "level", "preconditions", "postconditions", "extensions",
    "assumptions", "additional information", "solution hints", "dependencies", "limitations", "revision history")

ucs.descriptions <- ucs %>% 
    filter(Form == "description") %>% 
    mutate(
        Procedural = (rbinom(1, 1, 0.95)==1),
        Fields = paste0(c("title", "main scenario", sample(fields, rbinom(1, length(fields), 0.5))), collapse= "; ")
    )
```

### Procedural Use Case Descriptions

For all procedural use case descriptions, mock the values that will be used in the subsequent analyses.

```{r attributes}
ucs.procedural <- ucs.descriptions %>% 
    filter(Procedural) %>% 
    rowwise() %>% 
    mutate(
        Location = sample(c("Business Background", "Business Requirements", "Solution Proposal", "default"), 1),
        Level = sample(c("User", "System", "Business", "(no level)"), 1),

        # data obtained from the manual data extraction of quality attributes
        Text.Order = (rbinom(1, 1, 0.95)==1),
        Dependencies = (rbinom(1, 1, 0.9)==1),
        Coherent = runif(1),
        Steps = rbinom(1, 8+complexity*2, 0.6),
        Functional.Steps = rbinom(1, Steps, 0.8),
        Whitebox.Steps = rbinom(1, Functional.Steps, 0.1+0.3*(Location == "Solution Proposal")),
        Misplaced.Variations = rbinom(1, Functional.Steps, 0.2),
        Misplaced.Main = rbinom(1, Functional.Steps, 0.1),
        Sequence.Main = (rbinom(1, 1, 0.98)==1),
        Sequence.Alt = (rbinom(1, 1, 0.98)==1),
        Numbering = (rbinom(1, 1, 0.98)==1),
        Consistent.Grammar = rbinom(1, Steps, 0.1),

        # data obtained from the automatic data extraction
        entities = rbinom(1, 3, 0.3),
        actors = rbinom(1, entities, 0.5),
        actors.explicit = rbinom(1, actors, 0.3),
        interactions = rbinom(1, Steps*2, 0.6),
        pureinteractions = rbinom(1, interactions, 0.3+actors*0.2),
        consecutive = rbinom(1, interactions, 0.6)
    )
```

## Finalize

Assemble all data frames again

```{r assembly}
ucs <- merge(ucs, ucs.descriptions, all.x = TRUE)
ucs <- merge(ucs, ucs.procedural, all.x = TRUE)
```

Finally, export and save the mocked file.

```{r export}
write.csv(ucs, 'data/output/uca-annotations.csv')
```
