```{r setup, include=FALSE}
# cluster of libraries for easier R syntax
library(tidyverse)

# library for Bayesian modeling
library(brms)
library(ggdag)
```

# Inferential Analysis: Causes of Number of Use Cases

This notebook contains the inferential analysis about the causes of use case description quality.
In particular, we investigate which factors influence the number of use cases in a requirement.

The analysis consists of three steps according to the framework for statistical causal inference by Siebert:[^1]

1. Modeling: explicit visualization of causal assumptions
2. Identification: selection of variables from the causal model to involve in the statistical model
3. Estimation: implementation of a regression model to detect statistically significant effects.

```{r figdir}
figdir <- 'figures/inferential/causes/nuc/'
```

## Modeling

We start by making our causal assumptions explicit.
The following directed, acyclic graph represents a set of variables connected by edges where we assume a causal relationship.

```{r dag}
dag <- dagify(
        nuc ~ cplx + owner + author,

        exposure = "cplx",
        outcome = "nuc",

        labels = c(owner = "requirements owner", author = "requirements author", cplx = "complexity", nuc = "number of use cases"),
        coords = list(x = c(cplx = 0, owner = 0, author = 0, nuc = 1),
                        y = c(cplx = 1, nuc = 1, author = 0, owner = 2))
)

ggdag_status(dag.size, use_labels="label", text = FALSE) +
  guides(fill = "none", color = "none") + 
  theme_dag()
```

## Identification

To deconfound the effect of the exposure on the outcome variable, we need to identify the adjustment set, i.e., the set of variables to control.

```{r identification}
ggdag_adjustment_set(dag, exposure = "cplx", outcome = "nuc") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank(), 
        axis.text.y = element_blank(), axis.ticks.y = element_blank(), axis.title.y = element_blank())
```

## Estimation

In the third step, we estimate the average causal effect (ACE) of the exposure on the outcome.

### Formula and Distribution

The following regression formula represents our statistical model derived from the causal model.

```{r formula}
f <- bf(nuc ~ cplx + (1|owner) + (1|author))
```

### Data Preparation

```{r data-loading}
d <- read.csv('data/output/uca-aggregated.csv')
d$cplx <- factor(d$cplx, levels=c("Low Complexity", "Medium Complexity", "High Complexity"), ordered=TRUE)
```

We filter out all incomplete data points.

```{r filter-complete}
d  <- d[complete.cases(d$cplx), ]
d  <- d[complete.cases(d$owner), ]
d  <- d[complete.cases(d$author), ]

d <- d  %>% filter(nuc > 0 & form == "description" & procedural > 0)
```

### Priors

We need to assign a prior distribution to the coefficient of each predictor involved in the statistical model.

```{r eligible-priors}
get_prior(f, family = poisson, data = d)
```

We select uninformative priors for each coefficient.

```{r set-priors}
priors <- c(
    prior(normal(0, 1), class = Intercept),
    prior(normal(0, 1), class = b),
    prior(exponential(1), class = sd)
)
```

To confirm the eligibility of the priors, we run a prior predictive check.
To this end, we sample from the model without updating the coefficients based on the data.

```{r model-prior}
m.prior <-
  brm(data = d, family = poisson, f, prior = priors,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4, sample_prior="only",
    file = "./src/analysis/fits/cause.nuc.prior"
  )
```

Then, we draw 100 random samples from the prior model to assess, whether the priors are proper.

```{r prior-predictive-check}
brms::pp_check(m.prior, ndraws = 100)
```

Note that we would typically aim to see that the samples $y_{rep}$ encompass the actually observed data $y$. 
This may not be the case for the mocked data.

### Model Training

With the priors confirmed, we can train the model, i.e., update the coefficients based on the observed data.

```{r model}
m <-
  brm(data = d, family = poisson, f, prior = priors,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4, 
    file = "./src/analysis/fits/cause.nuc"
  )
```

Similarly, we can draw 100 random samples from the trained model to confirm that the model has updated its priors accordingly.

```{r posterior-predictive-check}
brms::pp_check(m, ndraws = 100)
```

### Evaluation

Finally, we can evaluate the model and derive conclusions from it.

#### Path Coefficients

To evaluate the model, we can inspect the posterior distribution of the path coefficients.

```{r coefficients}
summary(m)
```

These can also be visualized in an `mcmc_plot`.
Factors where the confidence does not overlap with 0 can be considered significant.

```{r mcmc-plot}
mcmc_plot(m)
```

#### Marginal Plots

Additionally, we can plot the marginal effects of specific predictors on the response variable of interest.
These marginal effects represent the isolated effect of the predictor on the response variable while holding all other predictors constant at representative values.

```{r marginal-cplx}
c_eff <- conditional_effects(m, effects="cplx")
plot(c_eff, plot = FALSE)[[1]] +
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed") +
  theme(axis.title.y = element_blank()) + 
  labs(y = "Number of use cases") + 
  scale_x_discrete(limits=rev) +
  coord_flip()

#ggsave(paste0(figdir, 'marginal-cplx.pdf'), width=4, height=1)
```
